A Wait-free Queue with Polylogarithmic Step Complexity

### 2023-01-10 podc_paper/podc-submission.tex
editing proof; fixed minor bug in algorithm (off-by-one error)
add garbage collection code to algorithm.tex
working on proof; some edits to intro to save space
working on proof (mainly indexdequeue)
add garbage collection lemmas
add algorithm2
Finished draft of analysis section.  (For some reason, compiling on my laptop requires commenting out some \newtheorem commands in podc-submission.tex .)
minor edits
Editing GC
some editing of GC
Rewording to save some space
minor edits to GC code
minor edits; changed tournament tree to ordering tree
Updated code to make GC run node-by-node instead
separ

\usepackage{tikz-qtree}
\usepackage{algorithm}
\usepackage{algpseudocode}
\makeatletter
\renewcommand{\ALG@beginalgorithmic}{\footnotesize}
\makeatother
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{relsize}
\usepackage{enumitem}
\usepackage{bold-extra}
\usepackage{showkeys}
\renewcommand*\contentsname{Table of Contents}

\usepackage{multicol}
\setlength\columnsep{24pt}

\def\draft{1}  % set to 1 for draft (include comments, color edits), 0 for clean final copy (omit comments, edits in black)
\if\draft 1
\newcommand{\mycomment}[3]{{\color{#2}{[\bf{#1: #3}]}}}
\newcommand{\myedit}[2]{{\color{#1}{#2}}\normalcolor}
\newcommand{\rebuttaledit}[2]{{\color{#1}{#2}}\normalcolor}
\newcommand{\here}[1]{\bf{[[[#1]]]}}
%\usepackage{showkeys}
\else
\newcommand{\mycomment}[3]{}
\newcommand{\rebuttaledit}[2]{#2}
\newcommand{\myedit}[2]{#2}
\newcommand{\here}[1]{}  % \here's include comments to ourselves that readers should not see.
\fi

@@

\newcommand{\node}{node}
\newcommand{\nodes}{nodes}
\newcommand{\block}{block}
\newcommand{\blocks}{blocks}

@@

\newcommand{\var}[1]{\mbox{\it{#1}}}
\newcommand{\fld}[1]{\mbox{\it{#1}}}
\newcommand{\typ}[1]{\mbox{\sf #1}}
\newcommand{\ceil}[1]{\lceil #1 \rceil}
\newcommand{\nl}{\mbox{\sf null}}
\newcommand{\tr}{\mbox{\sf true}}
\newcommand{\opa}[2]{\op{#1}(\var{#2})}
\newcommand{\op}[1]{{\sf #1}}
\newcommand{\opemph}[1]{{\sf \bf{#1}}}
\newcommand{\com}{$\triangleright$}
\newcommand{\linecomment}[1]{\com\ #1}

\usepackage{amsmath,amsthm}
%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{observation}{Observation}
%\theoremstyle{definition}
%\newtheorem{definition}[theorem]{Definition}
\newtheorem{invariant}{Invariant}
%\newtheorem{proposition}[theorem]{Proposition}

@@

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Eric Ruppert}
\affiliation{%
  \institution{York University}
  \streetaddress{P.O. Box 1212}
  \city{Toronto}
  \state{Ontario}
  \country{Canada}
  \postcode{43017-6221}
}
\email{ruppert@cse.yorku.ca}
\orcid{1234-5678-9012}

\author{Hossein Naderibeni}
\email{hnaderi268@gmail.com}
\affiliation{%
  \institution{York University}
  \streetaddress{P.O. Box 1212}
  \city{Toronto}
  \state{Ontario}
  \country{Canada}
  \postcode{43017-6221}
}

@@

\begin{abstract}
We introduce a novel linearizable wait-free queue implementation using single-word 
CAS instructions.
Previous lock-free queues all take $\Omega(p)$ steps per operation in the worst case, 
where $p$ is the number of processes that can access the queue.
We achieve $O(\log^2 p +\log q)$ steps per operation, where $q$ is the size of the queue.
\here{mention space usage; update time bound if space-efficient version takes more time per op}
\end{abstract}

@@

@@

\appendix
\end{document}

>>>>

\usepackage{tikz-qtree}
\usepackage{algorithm}
\usepackage{algpseudocode}
\makeatletter
\renewcommand{\ALG@beginalgorithmic}{\footnotesize}
\makeatother
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{relsize}
\usepackage{enumitem}
\usepackage{bold-extra}
\usepackage{cleveref}
\usepackage{thmtools,thm-restate}

%\renewcommand*\contentsname{Table of Contents}

\usepackage{multicol}
\setlength\columnsep{24pt}
%\renewcommand{\baselinestretch}{2}

\numberwithin{equation}{section}

\def\draft{0}  % set to 1 for draft (include comments, color edits), 0 for clean final copy (omit comments, edits in black)
\if\draft 1
\usepackage{showkeys}
\newcommand{\mycomment}[3]{{\color{#2}{[\bf{#1: #3}]}}}
\newcommand{\myedit}[2]{{\color{#1}{#2}}\normalcolor}
\newcommand{\rebuttaledit}[2]{{\color{#1}{#2}}\normalcolor}
\newcommand{\here}[1]{\bf{[[[#1]]]}}
%\usepackage{showkeys}
\else
\newcommand{\mycomment}[3]{}
\newcommand{\rebuttaledit}[2]{#2}
\newcommand{\myedit}[2]{#2}
\newcommand{\here}[1]{}  % \here's include comments to ourselves that readers should not see.
\fi

@@

\newcommand{\node}{node}
\newcommand{\nodes}{nodes}
\newcommand{\block}{block}
\newcommand{\blocks}{blocks}
\newcommand{\ordering}{ordering}

@@

\newcommand{\hangbox}[1]{\parbox[t]{\dimexpr\linewidth-\algorithmicindent\relax}{%
    \setlength{\hangindent}{\algorithmicindent}%
    #1}}
\newcommand{\spac}{\medskip}
\newcommand{\assign}{\mbox{:=}}

\newcommand{\var}[1]{\mbox{\textnormal{\it{#1}}}}
\newcommand{\fld}[1]{\mbox{\it{#1}}}
\newcommand{\typ}[1]{\mbox{\sf #1}}
\newcommand{\ceil}[1]{\lceil #1 \rceil}
\newcommand{\nl}{\mbox{\sf null}}
\newcommand{\tr}{\mbox{\sf true}}
\newcommand{\fa}{\mbox{\sf false}}
\newcommand{\opa}[2]{\op{#1(\var{#2})}}
\newcommand{\op}[1]{{\sf #1}}
\newcommand{\opemph}[1]{{\sf \bf{#1}}}
\newcommand{\com}{$\triangleright$}
\newcommand{\linecomment}[1]{\com\ #1}
\newcommand{\gc}[1]{\color{blue}#1\color{black}}

\usepackage{amsmath,amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{mytheorem}[theorem]{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{invariant}[theorem]{Invariant}
\newtheorem{claim}{Claim}[theorem]

\newtheorem{innercustomlemma}{Lemma}
\newenvironment{customlemma}[1]
  {\renewcommand\theinnercustomlemma{#1}\innercustomlemma}
  {\endinnercustomlemma}

@@

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

\author{Hossein Naderibeni}
\email{hnaderi268@gmail.com}
\affiliation{%
  \institution{York University}
  \streetaddress{P.O. Box 1212}
  \city{Toronto}
  \state{Ontario}
  \country{Canada}
  \postcode{43017-6221}
}
\author{Eric Ruppert}
\affiliation{%
  \institution{York University}
  \streetaddress{P.O. Box 1212}
  \city{Toronto}
  \state{Ontario}
  \country{Canada}
  \postcode{43017-6221}
}
\email{ruppert@cse.yorku.ca}
\orcid{0000-0001-5613-8701}

@@

\begin{abstract}
We introduce a novel linearizable wait-free queue implementation using single-word 
CAS instructions.
Previous lock-free queues all have amortized step complexity of $\Omega(p)$ per operation in worst-case executions, 
where $p$ is the number of processes that can access the queue.
We present a new wait-free queue that takes $O(\log p)$ steps per enqueue and $O(\log^2 p +\log q)$ steps per dequeue, where $q$ is the size of the queue.
A bounded-space version of the implementation has amortized step complexity of $O(\log^2 p +\log q)$ per operation.
\end{abstract}

@@

%OMIT IN SUBMISSION FOR ANONYMITY
%\paragraph{Acknowledgements}
%We thank Franck van Breugel for his helpful comments.
%Funding was provided by the Natural Sciences and Engineering Research Council of Canada.

@@

\newpage
\appendix

\input{correctness-details}

\input{algorithm-GC}

\input{correctness-GC}

\end{document}
